The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2012-06-06 script/de/chapters/cryptomethods.tex
CT script: Started adding chap 1.1 and OTP in 2.2.4 (DE only, translation of committed revision 3495)

\begin{sloppypar}
Einen sehr guten Überblick zu den Verfahren bietet das Buch von Bruce
Schneier \cite{cm:Schneier1996}. Grundsätz"-lich unterscheidet man zwischen
symmetrischen und asymmetrischen Verfahren zur Verschlüsse"-lung.

Aber davor soll es ein paar Informationen dazu geben, wie man Sicherheit von
Krytoalgorithmen definiert.
\end{sloppypar}

% --------------------------------------------------------------------------
\section[Sicherheits-Definitionen]{Sicherheits-Definitionen}

,,,,
Es gibt ein beweisbar sicheres Verschlüsselungsverfahren, 
das \index{One-Time-Pad} {\em One-Time-Pad}. Dieses weist allerdings
einige praktische Nachteile auf (der verwendete Schlüssel muss zufällig
gewählt werden und mindestens so lang wie die zu schützende Nachricht sein),
so dass es außer in geschlossenen Umgebungen, zum Beispiel beim heißen
Draht zwischen Moskau und Washington, kaum eine Rolle spielt.\par \vskip + 3pt

Für alle anderen Verfahren gibt es (theoretische) Möglichkeiten, sie
zu brechen. Bei guten Verfahren sind diese jedoch so aufwändig, dass sie
praktisch nicht durchführbar sind und diese Verfahren als (praktisch)
sicher angesehen werden können.\par \vskip + 3pt

>>>>

Zuerst erklären wir, wie die Sicherheit von Kryptosystemen definiert wird.

% --------------------------------------------------------------------------
\hypertarget{cm_Section_Security_Definitions}{}
\section[Sicherheits-Definitionen]{Sicherheits-Definitionen}
\label{cm_Section_Security_Definitions}

Modern cryptography is heavily based on mathematical theory and computer
science practice. Cryptographic algorithms are designed around computational
hardness assumptions, making such algorithms hard to break in practice by any
adversary.

\vskip +3pt
Depending on the adversary's capabilities there are mainly two basic notations
of security distinguished in literature
(see e.g. {\em Contemporary Cryptography} \cite{cm:Oppliger2011}):

\begin{itemize}

\item {\bf Computational, conditional or practical security}\\
A cipher is computationally secure if it is theoretically possible to break such
a system but it is infeasible to do so by any known practical means. Theoretical
advances (e.g., improvements in integer factorization algorithms) and faster
computing technology require these solutions to be continually adapted.

Even using the best known algorithm for breaking it will require so much
resources (e.g., 1,000,000 years) that practically the cryptosystem is secure.

So this concept is based on assumptions of the adversary's limited computing
power and the current state of science.

\item {\bf Information-theoretical or unconditional security}\\
A cipher is unconditionally secure if it is secure no matter how much resources
(time, space) the attacker has. So even in the case where an adversary has
unlimited resources for solving a problem e.g. breaking a cipher. An adversary
has not sufficient information for doing that because he is unable to gain any
meaningful data from a ciphertext.

There exist information-theoretically secure schemes that provably cannot be
broken even with unlimited computing power -- an example is the
{\em One Time Pad}\index{One Time Pad} (OTP\index{OTP}).

Das das OTP ein Informations-theoretisch sicheres Verschlüsselungsverfahren
ist, leitet sich seine Sicherheit schon allein aus der Informationstheorie
ab -- und ist sicher, auch wenn der Angreifer unbegrenzte Rechenkapazitäten
hat. Dieses Verfahren weist allerdings einige praktische Nachteile auf (der
verwendete Schlüssel dar nur einmal verwendet werden, muss zufällig
gewählt werden und mindestens so lang sein wie die zu schützende Nachricht),
so dass es außer in geschlossenen Umgebungen, zum Beispiel beim heißen
Draht zwischen Moskau und Washington, kaum eine Rolle spielt.%\par \vskip + 3pt

\vskip +3pt
Two more concepts are sometimes used:

\begin{itemize}

\item {\bf Provable security}
This means that breaking such a cryptographic system is as difficult as
solving some supposedly difficult problem e.g. discrete logarithm computation,
discrete square root computation, very large integer factorization.

Example: Currently we know that RSA is at most as difficult as factorization,
but we cannot prove that its exactly as difficult as factorization. So RSA
has no proven minimum security. Or in other words: We cannot prove, that if
RSA (the cryptosystem) is broken, that then factorization (the hard
mathematical problem) can be solved.

The Rabin cryptosystem was the first cryptosystem which could be proven to be
computationally equivalent to a hard problem.

\item {\bf Ad-hoc security}
A cryptographic system has this security feature if it is not worth to try to
break into such a system because of inadequate price of data with comparison to
price of work needed to do so. Or an attack can't be done in sufficiently short
time (see {\em Handbook of Applied Cryptography} \cite{cm:Menezes2001}).

Example: This may apply if a message relevant for the stock market will be
published tomorrow and you need a year to break it.

\vskip +3pt
Bei den heutzutage verwendeten guten Verfahren ist der Zeitaufwand zum Brechen
so hoch, dass sie praktisch nicht gebrochen werden können. Deshalb kann man
diese Verfahren als (praktisch) sicher ansehen -- aus einer rein auf den
Algorithmus bezogenen Sichtweise.%\par \vskip + 3pt

\begin{sloppypar}
Einen guten Überblick über die verschiedenen Verfahren bietet das Buch von Bruce
Schneier \cite{cm:Schneier1996}. Grundsätz"-lich unterscheidet man zwischen
symmetrischen und asymmetrischen Verfahren zur Verschlüsse"-lung.
\end{sloppypar}

### 2012-06-06 script/de/chapters/paper_and_pencil.tex
CT script: Started adding chap 1.1 and OTP in 2.2.4 (DE only, translation of committed revision 3495)

\item {\bf Superposition}\index{Superposition}\index{Verschlüsselung!Superposition}
   \begin{itemize}
      \item {\bf Running-Key-Chiffre}: 
         Addition eines Schlüsseltextes (z.B. aus einem Buch) zum Klartext.
      \item {\bf Überlagerung mit einer Zahlenfolge}: 
         Eine Möglichkeit sind mathematische Folgen wie die Fibonacci-Zahlen.
   \end{itemize}

>>>>

\item {\bf Superposition}\index{Superposition}\index{Verschlüsselung!Superposition}
   \begin{itemize}
      \item {\bf Running-Key-Chiffre}: 
         Addition eines Schlüsseltextes (z.B. aus einem Buch) zum Klartext.

      \item {\bf Überlagerung mit einer Zahlenfolge}: 
         Eine Möglichkeit sind mathematische Folgen wie die Fibonacci-Zahlen.

       \item {\bf One-Time-Pad (OTP)}\footnote{%
          In CT1\index{CT1} kann man dieses Verfahren über den Menüeintrag {\bf 
          Ver-/Entschlüsseln \textbackslash{} Symmetrisch (klassisch) 
          \textbackslash{} Vernam / OTP} aufrufen.}\index{One-Time-Pad}\index{OTP}:
          Eine Folge von Bytes wird Byte-für-Byte mit dem Klartext XOR-verknüpft.
          Dies ist eine Verallgemeinerung des Vigen\`ere-Verfahrens, und es war
          das erste Informations-theoretisch sichere Schema (vgl.
          Kapitel~\ref{cm_Section_Security_Definitions}
          ``\nameref{cm_Section_Security_Definitions}'').
          Für diese Anforderung muss das Pad zufällig sein und es darf nur
          einmal benutzt werden (um jede Ähnlichkeit von Patterns aus dem
          Geheimtext zu eliminieren).\\
          Begründung: Gegeben sei Geheimtext C, Klartext P, Pad K, und zwei
          Klartexte, die jeweils mit dem gleichen Schlüssel verschlüsselt sind:
          C1 = P1 $\oplus$ K;  C2 = P2 $\oplus$ K; \\
          Damit ergibt sich C1 $\oplus$ C2 = (P1 $\oplus$ K) $\oplus$
                           (P2 $\oplus$ K) = P1 $\oplus$ P2. \\
          Diese Kombination kann die Klartexte durchsickern lassen\footnote{%
          In JCT\index{JavaCrypTool}\index{JCT} kann man über den Menüeintrag
          {\bf Visualisierungen \textbackslash{} Viterbi} mit der automatischen
          Kryptoanalyse von Running-Key-Geheimtexten unter Verwendung der
          Viterbi-Analyse spielen.}.
   \end{itemize}

### 2012-06-06 script/en/chapters/cryptomethods.tex
CT script: Started adding chap 1.1 and OTP in 2.2.4 (DE only, translation of committed revision 3495)

As the OTP is information-theoretically secure it derives its security
solely from information theory and is secure even with unlimited computing
power at the adversary's disposal. However, OTP has several
practical disadvantages (the key used must be randomly selected and must be at
least as long as the message being protected), which means that it is hardly
used except in closed environments such as for the hot wire between Moscow and
Washington.%\par \vskip + 3pt

@@

\vskip +3pt
For good procedures used today the time taken to break them is so long that
it is practically impossible to do so, and these procedures can therefore be
considered (practically) secure.%\par \vskip + 3pt

>>>>

As the OTP is information-theoretically secure it derives its security
solely from information theory and is secure even with unlimited computing
power at the adversary's disposal. However, OTP has several practical
disadvantages (the key used must be used only once, randomly selected and must
be at least as long as the message being protected), which means that it is
hardly used except in closed environments such as for the hot wire between
Moscow and Washington.%\par \vskip + 3pt

@@

\vskip +3pt
For good procedures used today the time taken to break them is so long that
it is practically impossible to do so, and these procedures can therefore be
considered (practically) secure -- from a pure algorithm's point of view.%\par \vskip + 3pt

### 2012-06-06 script/en/chapters/paper_and_pencil.tex
CT script: Started adding chap 1.1 and OTP in 2.2.4 (DE only, translation of committed revision 3495)

       \item {\bf One Time Pad (OTP)}\footnote{%
          CT1\index{CT1} offers this method under the menu
          {\bf Encrypt/Decrypt \textbackslash{} Symmetric (classic)
          \textbackslash{} Vernam / OTP}.}\index{One Time Pad}\index{OTP}:
          A sequence
          of bytes is XORed byte-by-byte to the plaintext. This is a
          generalization of Vigen\`ere's mechanism and it was the first
          information-theoretically secure scheme (see
          chapter~\ref{cm_Section_Security_Definitions}
          ``\nameref{cm_Section_Security_Definitions}'').
          To fulfill this claim the pad must be random and it must be used
          only once (to eliminate any semblance of pattern from ciphertext).\\
          Reason: Given ciphertext C, plaintext P, pad K and two plaintexts
          encrypted with the same key: C1 = P1 + K;  C2 = P2 + K; \\
          Thus, C1 + C2 = (P1 + K) + (P2 + K) = P1 + P2; \\
          which effectively could leak the plaintexts\footnote{%
          In JCT\index{JavaCrypTool}\index{JCT} you can play with an automatic
          cryptanalysis of running-key ciphertexts using the Viterbi analysis
          under the menu {\bf Visuals \textbackslash{} Viterbi}.}.
   \end{itemize}

>>>>

       \item {\bf One Time Pad (OTP)}\footnote{%
          CT1\index{CT1} offers this method under the menu
          {\bf Encrypt/Decrypt \textbackslash{} Symmetric (classic)
          \textbackslash{} Vernam / OTP}.}\index{One Time Pad}\index{OTP}:
          A sequence
          of bytes is XORed byte-by-byte to the plaintext. This is a
          generalization of Vigen\`ere's mechanism and it was the first
          information-theoretically secure scheme (see
          chapter~\ref{cm_Section_Security_Definitions}
          ``\nameref{cm_Section_Security_Definitions}'').
          To fulfill this claim the pad must be random and it must be used
          only once (to eliminate any semblance of pattern from the ciphertext).\\
          Reason: Given ciphertext C, plaintext P, pad K and two plaintexts
          encrypted with the same key: C1 = P1 $\oplus$ K;  C2 = P2 $\oplus$ K; \\
          Thus, C1 $\oplus$ C2 = (P1 $\oplus$ K) $\oplus$ (P2 $\oplus$ K) = P1 $\oplus$ P2; \\
          which effectively could leak the plaintexts\footnote{%
          In JCT\index{JavaCrypTool}\index{JCT} you can play with an automatic
          cryptanalysis of running-key ciphertexts using the Viterbi analysis
          under the menu {\bf Visuals \textbackslash{} Viterbi}.}.
   \end{itemize}
